---
title: "Percentage of colocalizing mCherry-tagged proteins with Ede1-EGFP nuages in 3∆ background"
date: "Last compiled on `r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    code_download: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE,
                      dpi = 96, fig.width = 6, fig.height = 4)
```

```{r libs}
library(tidyverse)
library(Hmisc) # mean_sdl and binomial confs
library(ggbeeswarm)
```

# {.tabset .tabset-pills}

## Experiment


### Rationale

The aim of the experiment 
was to determine whether mCherry-tagged endocytic proteins
co-localize with Ede1-EGFP marked droplets
in *apl3∆ yap1801∆ yap1802∆* yeast cells. 

### Imaging and analysis

Two-color single planes were
acquired on the Olympus IX81
equipped with a 100x/1.49 objective.

Measuring colocalization by correlation
did not prove not very useful when applied to entire cells,
since all endocytic proteins 
are correlated in intensity in major regions
(background, cytoplasm and genuine endocytic patches).

I decided that a manual assessment would be more trustworthy.
Regions of interest containing Ede1-EGFP droplets
were cropped out of both channel images
based *only* on the examination of the GFP channel.
I used the ImageJ macro `blindAnalysis` to sequentially display all images
at random, blinding me to the strain of origin,
and recording a binary decision about colocalization
for each image.

### Strain list

```{r strains}
strains <- read_csv("data/raw/strain_lut.csv", 
                    col_types = cols(target  = col_factor(NULL))) # order factor levels per csv file
knitr::kable(strains)
```

## Results {.tabset}

```{r source_data, cache=TRUE}
files <- list.files(path = 'data/raw/', pattern = 'dataset',
                    full.names = TRUE)
counts <- map(files, read_csv,
              col_names = c('cell', 'colocalizes', 'dataset'),
              cols(dataset = col_factor(NULL))) %>%
          bind_rows() %>% # combine datasets
          separate(cell, c('date', 'strain', 'stack', 
                   'channel', 'processing', 'window'), '_') %>%
          left_join(strains, by = 'strain') %>%
          select(strain, stack, window, target, colocalizes, dataset)

write_excel_csv(format.data.frame(counts, digits = 3), 'data/colocalisation_data.csv')
save(counts, file = 'data/colocalisation_data.RData')
rm(list = ls()) # clean workspace
```

### Individual datasets

Percentage of colocalizing condensate for each target protein in each dataset.

x: number of colocalisations,
n: total number of analyzed condensates,
p: x / n, 
p_min and p_max: estimates of binomial confidence interval
based on sample size, by Wilson's method.

```{r load}
load('data/colocalisation_data.RData')
```

```{r individual_datasets}
proc <- counts %>%
  group_by(target, dataset) %>%
  summarise(x = sum(colocalizes), n = n(),
            p = binconf(x, n, method = 'wilson')[1],
            p_min = binconf(x, n, method = 'wilson')[2],
            p_max = binconf(x, n, method = 'wilson')[3])

knitr::kable(proc)

write_excel_csv(format.data.frame(proc, digits = 3), 'data/colocalisation_summary.csv')
```

### Averaged datasets

Percentage of colocalizing condensates for each target protein,
arithmetic mean ± SD from 3 datasets.

```{r averaged_datasets}
proc_stats <- proc %>%
  select(target, dataset, p) %>%
  group_by(target) %>%
  summarise(mean = mean(p), sd = sd(p), n = n())

knitr::kable(proc_stats)
```

## Plot {.active}

```{r}
theme_frap <- function(base_size = 11, base_family = "",
                       base_line_size = base_size / 22,
                       base_rect_size = base_size / 22) {
  # minimal theme with border
  # based on theme_linedraw without the grid lines
  # also trying to remove all backgrounds but plot_background won't set to blank
  theme_linedraw(
    base_size = base_size,
    base_family = base_family,
    base_line_size = base_line_size,
    base_rect_size = base_rect_size
  ) %+replace%
    theme(
      # no grid and no backgrounds if I can help it
      plot.background = element_blank(),
      panel.background = element_blank(),
      panel.grid = element_blank(),
      plot.margin = margin(0, 0, 0, 0),
      complete = TRUE
    )
}

myplot <- proc %>% ggplot(aes(x = target, y = p)) + 
  geom_quasirandom(aes(shape = dataset), fill = 'gray80',
              size = 2, width = 0.3, show.legend = FALSE)+
  stat_summary(fun.y = mean, geom = "errorbar", 
               aes(ymax = ..y.., ymin = ..y..),
               width = 0.5, size = 1)+
  stat_summary(fun.data = 'mean_sdl', geom = 'errorbar',
               fun.args = list(mult = 1), width = 0.2)+
  #scale_color_brewer(palette = 'Set2')+
  scale_shape_manual(values = c(21:25))+ # points with borders
  theme_frap(base_size = 18, base_family = "Myriad Pro")+
  scale_y_continuous(labels = scales::label_percent(suffix = ''))+
  theme(legend.position = c(0.95, 0.95),
        legend.justification = c("right", "top"),
        legend.background = element_rect(fill = NULL, colour = 'black',
                                         size = 0.5))+
  labs(title = NULL,
       y = 'Fraction of colocalizing condensates (%)',
       x = 'Target protein')

print(myplot)
ggsave('figures/colocalizations.pdf', device = cairo_pdf, width = 160, height = 120, units = "mm")
```

## Source data

### .csv

```{r echo=FALSE}
xfun::embed_file('data/colocalisation_data.csv')
```

### .RData

```{r echo=FALSE}
xfun::embed_file('data/colocalisation_data.RData')
```

## Session info

```{r session, message=TRUE}
sessionInfo()
```
