---
title: "Sla1 patch density in Ede1∆PQCC-Syp1 fusion"
date: "Last compiled on `r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    code_download: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE,
                      dpi = 96, fig.width = 4, fig.height = 3)
```

```{r libs}
library(tidyverse) 
library(broom) # to summarise significance tests in a tidy manner
library(ggbeeswarm) # beeswarm scatter plots
library(ggsignif) # significance on plots
library(knitr)
library(rstatix)
```

```{r load}
rm(list = ls())
load("data/syp1_fusion_sla1.RData")
```

```{r theme}
# Custom ggplot2 theme
# --------------------

# minimal theme with border
# based on theme_linedraw without the grid lines
# also trying to remove all backgrounds and margins
# the aim is to make it as easy as possible to edit in illustrator

theme_clean <- function(base_size = 11, base_family = "",
                        base_line_size = base_size / 22,
                        base_rect_size = base_size / 22) {
  theme_linedraw(
    base_size = base_size,
    base_family = base_family,
    base_line_size = base_line_size,
    base_rect_size = base_rect_size
  ) %+replace%
    theme(
      # no grid and no backgrounds if I can help it
      legend.background =  element_blank(),
      panel.background = element_blank(),
      panel.grid = element_blank(),
      plot.background = element_blank(),
      plot.margin = margin(0, 0, 0, 0),
      complete = TRUE
    )
}

# Set default theme
# -----------------
theme_set(theme_clean(base_size = 14, base_family = "Myriad Pro"))

# Create a ggsave wrapper
# -----------------------

# This way we can set a default size and device for all plots
my_ggsave <- function(filename, plot = last_plot(),
                      device = cairo_pdf, units = "mm",
                      width = 100, height = 80, ...){
  ggsave(filename = filename, plot = plot,
         device = device, units = units,
         height = height, width = width,  ...)
  }
```

# {.tabset .tabset-pills}

## Experiment

The aim of the experiment is to determine Sla1 patch density where the central region deletion (PQ-rich and coiled-coil) of Ede1 is attached to Syp1
using the FRB-FKBP system.

### Imaging

All images were captured with the halide lamp @ 50% power
using the GFP filter cube and 400 ms exposure per slice.
5 micron stacks were taken with 0.2 micron step.

### List of strains used

Strain | Mutations
--- | ---
MKY0140 | wild type
MKY0654 | ede1∆
MKY4343 | Ede1∆(366-900)-mCherry-FKBP, SYP1-FRB, fpr1∆, tor1-1

All: Sla1-EGFP

MKY4343 was cultured with 10 ug/ml rapamycin (+R) or equivalent volume of solvent
(1% DMSO in culture, -R).

### Image processing

Patch numbers were extracted with a custom Python script `site_counter.py` 
(part of package impromptu called `mkimage`)
using functions from the `scikit-image` package. 
Briefly, the images were median-filter subtracted with a 5 px disk brush
and thresholded using the Yen method.
The thresholded image was eroded using the number of non-zero
neighbouring pixels in 3D as the erosion criterion.
The spots were counted using skimage.measure.label() function with 2-connectivity.

Cross-section area was obtained by median-filtering of the stack
with 10px disk brush, calculating the maximum projection image, 
thresholding using Otsu's algorithm and using skimage.measure.regionprops()
to obtain area.
Note that these are pixel counts of *cross-section* area. 
To determine the total surface area, I assume that an unbudded cell is spherical
(suface area is four times the cross-section area).

This call to `site_counter` was used to process all datasets:

```
process_folder(path, median_radius = 5, erosion_n = 1, con = 2,
                   method = Yen, mask = False, loop = False, save_images = True)
```

`data_cleanup.Rmd` was used to gather all output into tidy data frames
with no further modifications.

## Statistical summary {.tabset}

### Sla1 density

#### Per dataset

We measured the number of patches and cell surface area.
We can combine these these as $density = \frac{patches}{area}$,
calculated for each cell individually.

Data summary for each Ede1 mutant in each dataset:

```{r density}
sla1_density_stats <- sla1_density %>%
  group_by(ede1, dataset) %>%
  summarise(n = n(),
            across(density,
                   list(mean = mean, sd = sd, median = median, mad = mad,
                        shapiro.p = ~ tidy(shapiro.test(.x))$p.value)),
            .groups = 'drop')

sla1_density_stats %>% kable()
```

#### For all repeats

Summary statistics for all experiments, derived from *mean values*
of N independent repeats.

```{r}
sla1_density_stats %>%
  group_by(ede1)%>%
  summarise(N = n(),
            across(density_mean,
                   list(mean = mean, sd = sd, median = median, mad = mad),
                        #shapiro.p = ~ tidy(shapiro.test(.x))$p.value),
                       .names = '{.fn}'),
            .groups = 'drop') %>%
  kable()
```

### Hypothesis testing

#### Are the Ede1 groups significantly different?

We will use a linear model of `density_mean ~ ede1`
to perform an ANOVA test on replicate-level means, 
followed by a Tukey test for multiple comparisons.

```{r}
anova <- anova_test(sla1_density_stats, density_mean ~ ede1)

anova %>%
  kable()
```

One-way ANOVA rejects the null, followed by Tukey's range test:

```{r}
tukey <- tukey_hsd(sla1_density_stats, density_mean ~ ede1,
                   ordered = TRUE)

tukey %>%
  kable()
```

The differences between DMSO control and rapamycin treatment
are not significantly different.

## Plots {.tabset}

```{r density_scatter}
plot_blank <- ggplot(sla1_density_stats,
                     aes(x = ede1, y = density_mean))+
  labs(title = NULL, x = NULL, y = expression("Sla1 patches/µm"^2))+
  scale_y_continuous(breaks = seq(0.0, 1.0, 0.1))+
  scale_shape_manual(values = c(21:25))+
  scale_color_brewer(palette = 'Pastel2')+
  scale_fill_brewer(palette = 'Set2')

plot_scatter <- plot_blank + 
  geom_quasirandom(data = sla1_density,
                   aes(x = ede1, y = density,
                     shape = dataset, #colour = dataset
                     ),
                 colour = 'grey75',
                 show.legend = F, size = 0.8
                 )

plot_violin <- plot_blank + 
  geom_violin(inherit.aes = F,
              data = sla1_density, aes(x = ede1, y = density),
              colour = 'grey70', fill = 'transparent'
              )
```

### SuperPlots

```{r}
plot_super <- plot_scatter +
  geom_quasirandom(aes(shape = dataset, fill = dataset),
                   show.legend = F,
                   width = 0.3, size = 2)+
  stat_summary(fun = mean, geom = 'crossbar',
               width = 0.5, fatten = 1)+
  stat_summary(fun.data = 'mean_sdl',
               fun.args = list(mult = 1), 
               geom = 'errorbar',
               width = 0.2)
  
print(plot_super)
my_ggsave('figures/density_super.pdf')

plot_super_violin <- plot_violin +
  geom_quasirandom(aes(shape = dataset, fill = dataset),
                   show.legend = F,
                   width = 0.3, size = 2)+
  stat_summary(fun = mean, geom = 'crossbar',
               width = 0.5, fatten = 1)+
  stat_summary(fun.data = 'mean_sdl',
               fun.args = list(mult = 1), 
               geom = 'errorbar',
               width = 0.2)

print(plot_super_violin)
my_ggsave('figures/density_super_violin.pdf')
```

### SuperPlots with significance

Let's add significance stars based on Tukey's test.

```{r}
extract_comparisons <- function(x, rows){
  #' Extract comparisons from rstatix tidy tests
  #' 
  #' This function subsets the selected comparisons in a Tukey,
  #' Dunn or similar test done by rstatix.
  #' It converts groups to a list of vectors that can be passed to geom_signif
  #'
  #'
  #' @param x: df or tibble, the comparison results
  #' @param rows: integer vector, the rows with desired comparisons
  x_subset <- x[rows,] %>%
    .[nrow(.):1,]
  
  x_comparisons <- x_subset %>%
    select(group1, group2) %>%
    t() %>%
    as.data.frame() %>%
    as.list
  x_annotations <- x_subset$p.adj.signif %>%
    as.vector()
  
  significance <- list(
    comparisons = x_comparisons,
    annotations = x_annotations
  )
  
  return(significance)
}
```

```{r}
significance <- extract_comparisons(tukey, c(1:6))

plot_super_signif <- plot_super +
  geom_signif(comparisons = significance$comparisons,
              annotations = significance$annotations,
              step_increase = 0.03,
              tip_length = 0.01, vjust = 0.8,
              margin_top = -0.1)
print(plot_super_signif)
my_ggsave('figures/density_super_signif.pdf')
```

#### Letter annotations

In this view, groups sharing at least one letter are not significant
at a chosen $\alpha$ (here, 5%).

Pros:

  * less cluttered
  * simpler to read and edit

Cons:

  * cannot distinguish different confidence levels
  * focus on non-findings

```{r}
library(multcompView)

add_tukey_labels <- function(x, yvar, xvar, alpha = 0.95){
  #' Add letter group labels generated by Tukey's HSD
  #'
  #' This function performs ANOVA and Tukey's HSD,
  #' extracts the letter labels and attaches them
  #' to the original data.
  #'
  #' @param x: df or tibble, the data (long format!)
  #' @param yvar: chr, name of the dependent variable
  #' @param xvar: chr, name of the independent variable
  #' @param alpha: dbl, confidence level passed on to Tukey's test
  #'
  aov_form <- formula(paste(yvar, '~', xvar))
  anova <- aov(formula = aov_form, data = x)
  tukey <- TukeyHSD(anova, which = xvar, conf.level = alpha)
  # Extract labels and factor levels from Tukey post-hoc 
  x_labels <- as_tibble(
    multcompLetters4(anova, tukey)[[xvar]]$Letters,
    rownames = xvar
    )
  x_labels <- rename(x_labels, tukey_group = value)
  x <- left_join(x, x_labels, by = xvar)
  
  return(x)
  }
``` 

```{r}
sla1_density_stats <- sla1_density_stats %>%
  mutate(ede1 = recode_factor(ede1, '-R' = 'noR', '+R' = 'R')) %>%
  add_tukey_labels('density_mean', 'ede1') %>%
  mutate(ede1 = recode_factor(ede1, 'noR' = '-R', 'R' = '+R'))
```

```{r}
plot_super_letters <- plot_super +
  geom_text(data = sla1_density_stats,
            aes(label = tukey_group), y = Inf, vjust = 1.2)
  #stat_summary(aes(label = tukey_group),
  #             fun.y = Inf,
  #             geom = 'text', na.rm = T, vjust = -0.5,
  #             #fun.args = list(err = errors, mult = error_range)
  #             )
print(plot_super_letters)
my_ggsave('figures/density_super_letters.pdf')
```


## Source data

### .csv

```{r echo=FALSE}
xfun::embed_file('data/syp1_fusion_sla1.csv')
```

### .RData

```{r echo=FALSE}
xfun::embed_file('data/syp1_fusion_sla1.RData')
```

## Session info

```{r session, message=TRUE}
sessionInfo()
```
