---
title: "Percentage of nuage-forming cells in 3∆ background Ede1 truncation mutants"
date: "Last compiled on `r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    code_download: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE,
                      dpi = 96)
```

```{r libs}
library(tidyverse)
library(Hmisc) # mean_sdl and binomial confs
library(ggbeeswarm)
```

# {.tabset}

## Experiment

### Rationale

I want to quantify the nuage-forming tendencies
of each Ede1 truncation construct,
in two yeast cell backgrounds (wild-type and *apl3∆ yap1801∆ yap1802∆*)
Absent a more finely tuned method, 
I have counted the percentage of condensate-containing cells
(by subjective but blinded assessment).

### Method

I acquired all data on the Olympus IX81
equipped with a 100x/1.49 objective, 
using the X-Cite 120PC lamp at 50% intensity 
and 400 ms exposure for illumination. 
Light was filtered through a U-MGFPHQ filter cube. 
I acquired stacks of 26 planes with a step size of 0.2 microns.

Individual cells were cropped out of fields of view
based only on the bright field channel.
I used the ImageJ macro `blindAnalysis` 
to sequentially display all images
at random, blinding me to the strain of origin,
and recording a binary decision about condensate formation
for each image.

### Strain list

```{r strains}
strains <- read_csv("data/input/strain_lut.csv", 
                    col_types = cols(bg = col_factor(NULL), Ede1 = col_factor(NULL))) # order factor levels in csv file
knitr::kable(strains)
```

```{r source_data}
# initialize a data frame with counts and get information about experiment from file names
files <- list.files(path = 'data/input/', pattern = 'dataset',
                    full.names = TRUE)

counts <- map(files, read_csv,
              col_names = c('cell', 'nuages', 'dataset'),
              cols(dataset = col_factor(NULL))) %>%
          bind_rows() %>%
          # separate filename into variables
          separate(cell, c('date', 'strain',
                           'stack', 'channel', 'cell'), '_') %>%
          left_join(strains, by = 'strain') %>%
          select(strain, stack, cell, Ede1, nuages, bg, dataset)

write_excel_csv(format.data.frame(counts, digits = 3), 'data/truncation_data.csv')
save(counts, file = 'data/truncation_data.RData')
rm(list = ls()) # clean workspace
```

```{r load}
load('data/truncation_data.RData')
```

## Data summary {.tabset}

### Individual datasets

Percentage of cells with condensates for each Ede1 construct in each background, by dataset.

x: number of cells with condensates, n: total number of analyzed condensates, p: x / n, p_min and p_max: estimates of binomial confidence interval based on sample size, by Wilson's method.

```{r}
proc <- counts %>%
  group_by(bg, Ede1, dataset) %>%
  summarise(x = sum(nuages), n = n(),
            p = binconf(x, n, method = 'wilson')[1],
            p_min = binconf(x, n, method = 'wilson')[2],
            p_max = binconf(x, n, method = 'wilson')[3])

knitr::kable(proc)
write_excel_csv(format.data.frame(proc, digits = 3), 'data/truncation_summary.csv')
```

### Averaged datasets {.active}

```{r averaged_datasets}
proc_stats <- proc %>%
  select(bg, Ede1, dataset, p) %>%
  group_by(bg, Ede1) %>%
  summarise(mean = mean(p), sd = sd(p), n = n())

knitr::kable(proc_stats)
```

## Plots {.tabset}

```{r}
theme_frap <- function(base_size = 11, base_family = "",
                       base_line_size = base_size / 22,
                       base_rect_size = base_size / 22) {
  # minimal theme with border
  # based on theme_linedraw without the grid lines
  # also trying to remove all backgrounds but plot_background won't set to blank
  theme_linedraw(
    base_size = base_size,
    base_family = base_family,
    base_line_size = base_line_size,
    base_rect_size = base_rect_size
  ) %+replace%
    theme(
      # no grid and no backgrounds if I can help it
      plot.background = element_blank(),
      panel.background = element_blank(),
      panel.grid = element_blank(),
      
      complete = TRUE
    )
}
```

### Lineplot

This is a difficult plot because of the number of variables involved: Ede1 truncation, strain background & dataset. First all on one plot, and means as thick lines.

```{r}
plot <- ggplot(proc, aes(x = Ede1, y = p, group = bg, fill = bg)) + 
  #geom_bar(stat = "identity", position = 'dodge', width = 0.5, colour = 'black')+
  #scale_fill_grey(start = 1, end = 0.5)+
  #geom_errorbar(aes(ymin = p_min, ymax = p_max), position = position_dodge(width = 0.5), width = 0.2) +
  stat_summary(fun.y = mean, geom = "errorbar",
               position = position_dodge( width = 0.7),
               aes(ymax = ..y.., ymin = ..y..),
               width = 0.5, size = 1)+
  stat_summary(fun.data = 'mean_sdl', geom = 'errorbar',
               position = position_dodge(width = 0.7),
               fun.args = list(mult = 1), width = 0.2)+
  geom_point(#aes(shape = dataset),
             shape = 21,
             position = position_jitterdodge(),
             size = 2, width = 0.3)+
  scale_color_brewer(palette = 'Set2')+
  scale_shape_manual(values = c(21:25))+
  theme_frap(base_size = 16, base_family = "Myriad Pro")+
  scale_y_continuous(labels = scales::label_percent(suffix = ''))+
  theme(legend.position = c(0.9, 0.5),
        legend.background = element_rect(fill = NULL, colour = 'black',
                                         size = 0.5))+
  #theme(legend.position = "none")+
  #geom_text(aes(label = scales::percent(proc)), stat= "identity", vjust = -0.5) +
  #geom_text(aes(label = total), stat= "identity", vjust = -0.5, hjust = 'middle', position = position_dodge(width = .5)) +
  #facet_grid(rows = vars(bg))+
  labs(title = NULL,
       y = 'Cells with condensates (%)')

print(plot)
ggsave('figures/nuages.pdf', device = cairo_pdf, width = 160, height = 120, unit = 'mm')
```

### Bar graph {.active}

The lines here do not provide enough visual cues as to which mean and point belongs to which combination of groups. As disliked as bar graphs are, I think in this case they would be much better suited to guide the reader.

```{r}
plot <- ggplot(proc, aes(x = Ede1, y = p, group = bg, fill = bg)) + 

  stat_summary(fun.y = mean, geom = "bar",
               position = position_dodge(),
               width = 0.5, colour = 'black')+
  stat_summary(fun.data = 'mean_sdl', geom = 'errorbar',
               position = position_dodge(width = 0.5),
               fun.args = list(mult = 1), width = 0.3)+
  geom_point(aes(shape = dataset),
             position = position_jitterdodge(jitter.width = 0.4, 
                                             dodge.width = 0.5),
             size = 2, show.legend = FALSE)+
  scale_color_brewer(palette = 'Set2')+
  scale_fill_grey(start = 1, end = 0.5)+
  scale_shape_manual(values = c(21:25))+
  theme_frap(base_size = 16, base_family = "Myriad Pro")+
  scale_y_continuous(labels = scales::label_percent(suffix = ''))+
  theme(legend.position = c(0.5, 0.95),
        legend.justification = c("center", "top"),
        #legend.box.just = "right",
        legend.margin = margin(0, 6, 6, 6),
        legend.background = element_rect(fill = NULL, colour = 'black',
                                         size = 0.5),
        legend.title = element_blank())+
  labs(title = NULL,
       y = 'Cells with condensates (%)')

print(plot)
ggsave('figures/nuages_bar.pdf', device = cairo_pdf, width = 160, height = 120, unit = 'mm')
```

### Faceted line plot

Finally, for the easily confused (by multiple groupings) and the easily offended (by bar graphs), go back to lines but facet the two backgrounds.

```{r}
plot <- ggplot(proc, aes(x = Ede1, y = p, group = bg, fill = bg)) + 
  stat_summary(fun.y = mean, geom = "errorbar",
               aes(ymax = ..y.., ymin = ..y..),
               width = 0.5, size = 1)+
  stat_summary(fun.data = 'mean_sdl', geom = 'errorbar',
               fun.args = list(mult = 1), width = 0.2)+
  geom_point(aes(shape = dataset, fill = dataset),
             position = position_jitter(width = 0.1),
             size = 2, show.legend = FALSE)+
  scale_color_brewer(palette = 'Set2')+
  scale_fill_grey(start = 1, end = 0.5)+
  scale_shape_manual(values = c(21:25))+
  theme_frap(base_size = 16, base_family = "Myriad Pro")+
  scale_y_continuous(labels = scales::label_percent(suffix = ''))+
  theme(legend.position = 'right',
        legend.title = element_blank(),
        axis.text.x = element_text(angle = -45, vjust = 0.1, hjust = 0.1))+
  # rotating axis labels: will be available when I update ggplot
  # I don't want to do it yet though as the potential fuckups are countless
  # guides(x =  guide_axis(angle = 90))+
  facet_grid(rows = vars(bg))+
  labs(title = NULL,
       y = 'Cells with condensates (%)')

print(plot)
ggsave('figures/nuages_facet.pdf', device = cairo_pdf, width = 120, height = 120, unit = 'mm')
```

## Source data

### .csv

```{r embed_csv}
xfun::embed_file('data/truncation_data.csv')
```

### .RData

```{r embed_rdata}
xfun::embed_file('data/truncation_data.RData')
```

## Session info

```{r session, message=TRUE}
sessionInfo()
```
