---
title: "Ede1 levels in 3∆ and OE strains"
date: "Last compiled on `r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    code_download: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE,
                      dpi = 96, fig.width =6, fig.height = 4)
```

```{r libs}
library(tidyverse)
library(knitr)
library(broom)

# ggplot2 extensions
library(ggbeeswarm)
library(ggsignif)

# for the linear mixed models
library(lmerTest)
library(emmeans)
```

```{r load}
# Load data generated by cleanup notebook
rm(list = ls())
load('data/data_clean.RData')
```

```{r theme}
# Custom ggplot2 theme
# --------------------

# minimal theme with border
# based on theme_linedraw without the grid lines
# also trying to remove all backgrounds and margins
# the aim is to make it as easy as possible to edit in illustrator

theme_clean <- function(base_size = 11, base_family = "",
                        base_line_size = base_size / 22,
                        base_rect_size = base_size / 22) {
  theme_linedraw(
    base_size = base_size,
    base_family = base_family,
    base_line_size = base_line_size,
    base_rect_size = base_rect_size
  ) %+replace%
    theme(
      # no grid and no backgrounds if I can help it
      legend.background =  element_blank(),
      panel.background = element_blank(),
      panel.grid = element_blank(),
      plot.background = element_blank(),
      plot.margin = margin(0, 0, 0, 0),
      complete = TRUE
    )
}

# Set default theme
# -----------------
theme_set(theme_clean(base_size = 14, base_family = "Myriad Pro"))

# Create a ggsave wrapper
# -----------------------

# This way we can set a default size and device for all plots
my_ggsave <- function(filename, plot = last_plot(),
                      device = cairo_pdf, units = "mm",
                      width = 100, height = 80, ...){
  ggsave(filename = filename, plot = plot,
         device = device, units = units,
         height = height, width = width,  ...)
  }
```

# {.tabset .tabset-pills}

## Experiment

### Premise

The aim of the experiment is to visualize the changes
in total and cytosolic concentrations of EGFP-tagged Ede1 
in the three strains used throughout the paper: 
  wild-type background, 
  3∆,
  and overexpression driven by the ADH1 promoter.

### Imaging

All strains within one dataset 
were imaged on the same day and in the same conditions 
on the Nikon Ti microscope
equipped with a Yokogawa CSW1 spinning disk
and a Photometrics Prime 95B sCMOS camera. 
The Cellular and Cytosolic acquisitions 
were taken on different days
so they are not directly comparable; 
the only meaningful comparison 
is between different strains within one localization / dataset combination.

Two different series of exposures were taken: 
high exposure for cytosolic quantification (saturated condensates)
and low exposure for total quantification (avoiding any saturation).
The cytosolic quantification was done on cells co-expressing Abp1-mCherry,
and the total on cells co-expressing Rvs167-mCherry (explanation below).

#### Strain list

```{r strains}
kable(strains)
```

#### Cytosolic intensity

Single planes were acquired at the equatorial plane using
  488 nm laser at 100% power, 
  with 500 ms exposure,
  and camera gain level 2.

The three strains used were the ones co-expressing Abp1-mCherry.

#### Total cellular intensity

Two-color stacks were acquired with 
  0.2 μm spacing
  and 5 μm range
  around the equatorial plane. 

* Ede1-EGFP: 488 nm laser at 
  15% power, 
  with 100 ms exposure,
  and camera gain level 2.
* Rvs167-mCherry: 561 nm laser at 
  100% power, with 
  500 ms exposure,
  and camera gain level 3.

Initially both the Cytosolic and Cellular datasets
were taken during the same acquisition, 
using the Abp1-mCherry strains.
However, Abp1 turned out suboptimal for masking entire cells 
(too much contrast between sites and cytosol),
and I repeated the experiment with the Rvs167 strains. 
The diffuse signal of Rvs167-mCherry 
makes it a perfect protein to create cellular masks. 
The results were ultimately very similar on average, 
but I choose to show Ede1-GFP/Rvs167-mCherry 
because the quantification is more precise.

### Processing

All images were background-subtracted using the ImageJ rolling ball algorithm 
with a 50px radius.

#### Cytosolic intensity
 
5x5 px square regions away from the condensates and vacuoles
were manually selected in ImageJ and mean pixel intensity was saved to file.

#### Total cellular intensity

Individual cells were cropped in ImageJ
and the function `batch_mask` from my personal Python package `mkimage`
was used to quantify intensity. 
Briefly, the RFP channel was 
  median-filtered with a disk brush of 10 px radius, 
  and thresholded using Li's method 
  followed by one round of morphological opening
to create a mask.
The mask was applied to GFP images and summary statistics 
of the masked regions (mean, median, sd) were saved.

#### Data clean-up

The mean pixel values of cytosolic regions and entire cells 
were loaded from text files using a separate R notebook. 
That notebook was used only to gather all observations 
in the tidy data structures accessible here, without any further processing.

## Data summary

```{r scale}
# generate normalized intensities (within each dataset)

# pull out scaling factor
wt_means <- intensities %>%
  filter(strain == 'wt') %>%
  group_by(dataset, localization) %>%
  summarise(wt_mean = mean(intensity), .groups = 'drop')

# join with the dataframe, generate new variable
intensities <- left_join(intensities, wt_means,
                         by = c('dataset', 'localization')) %>%
  mutate(normalized = intensity / wt_mean)
```

The table below summarizes intensity measurements from each dataset,
as well as normalized intensities. The normalization was performed for each
dataset by dividing each observation by the mean wild-type intensity
for that dataset.

```{r summarize_normal}
# generate summary statistics
intensity_stats <- intensities %>%
  group_by(localization, strain, dataset) %>%
  summarise(n = n(),
            across(c(intensity, normalized),
                   list(mean = mean, sd = sd, median = median,
                        se = ~ sd(.x) / sqrt(n())
                        )), .groups = 'drop')

intensity_stats %>%
  kable()
```

### Aggregated normalized means

Quick summary of `normalized mean` from above
expressed as average values of repeat-level means
with 95% confidence intervals:

```{r}
mean_ci <- intensity_stats %>%
  group_by(localization, strain)%>%
  summarise(round(mean_cl_normal(normalized_mean), 2)) %>%
  rename(mean = y, lower = ymin, upper = ymax)

kable(mean_ci)
```

This hints that the true mean of cytosolic fluorescence 
is unchanged in different strains,
and that 3Δ and overexpression strains
feature some (sgnificant?) levels of overexpression.
More accurate estimates and p-values can be found in the 'Modeling' tab.

## Plots 

I have chosen to show this data using the 
[SuperPlot](https://doi.org/10.1083/jcb.202001064) style.
Each point shows mean intensity of a small region in the cytoplasm
of an individual cell, or one entire cell (separated into facets).

Big colour points show mean measurements from three independent repeats.

Range is mean +/- SD, calculated based on the three independent repeats.

### Without normalization

These plots show raw intensity measures. The dataset effect is not so bad
in the Cellular data, which were imaged 3 days in a row. It is quite pronounced
in the Cytosolic data, where the facility realigned the laser fiber in the microscope 
between datasets 2 and 3, affecting the absolute values.

Also, if it seems weird that the cytosolic values are higher
than the cellular ones, that is not accidental. The exposure of the cellular
data had to be kept low in order to avoid saturating the condensates.
In the cytosolic data, the exposures were high in order to maximize
the cytosolic signal, allowing for the condensates to saturate.

```{r}
plot_base <- ggplot(intensity_stats,
                   aes(x = strain, y = intensity_mean)) +
  labs(title = NULL, x = 'Strain', y = 'Intensity (a. u.)')+
  scale_y_continuous(breaks = scales::breaks_extended(6))+
  scale_shape_manual(values = c(21:25)) +
  scale_color_brewer(palette = 'Set2') +
  scale_fill_brewer(palette = 'Set2') +
  facet_wrap(. ~localization,
             scales = 'free')


plot_scatter <- plot_base + 
  geom_quasirandom(inherit.aes = F, data = intensities,
                   aes(x = strain, y = intensity,
                       shape = dataset),
                   colour = 'gray75', #width = 0.3,
                   size = 0.8, #varwidth = TRUE,
                   show.legend = F
                   )

plot_violin <- plot_base + 
  geom_violin(inherit.aes = F, data = intensities,
              aes(x = strain, y = intensity),
              scale = 'width',
              colour = 'gray75', fill = 'transparent'
              )
```

```{r}
plot_super <- plot_scatter +
  geom_quasirandom(aes(shape = dataset, fill = dataset),
                   show.legend = F, width = 0.3, size = 2) +
  stat_summary(fun = mean, geom = "crossbar", 
               width = 0.5, fatten = 1) +
  stat_summary(fun.data = 'mean_sdl', geom = 'errorbar',
               fun.args = list(mult = 1), width = 0.2)
print(plot_super)
my_ggsave('figures/super_raw.pdf')
```

```{r}
# These significance bars are generated by `ggsignif` from pairwise t-tests.
# They are not entirely correct and so they are not shown in the notebook.
# The figure is saved for me, to manually change the annotations
# in Illustrator to the ones from multiple comparisons tests.

plot_signif <- plot_super + 
  geom_signif(comparisons = list(c("wt", "3Δ"),
                                 c("3Δ", "oe"),
                                 c("wt", "oe")),
              map_signif_level = T, test = 't.test',
              step_increase = 0.03, tip_length = 0.01, vjust = 0.8)

my_ggsave('figures/super_raw_signif.pdf', plot = plot_signif)
```


```{r}
plot_super_violin <- plot_violin +
  geom_quasirandom(aes(shape = dataset, fill = dataset),
                   show.legend = F, width = 0.3, size = 2) +
  stat_summary(fun = mean, geom = "crossbar", 
               width = 0.5, fatten = 1) +
  stat_summary(fun.data = 'mean_sdl', geom = 'errorbar',
               fun.args = list(mult = 1), width = 0.2)
#print(plot_super_violin)
my_ggsave('figures/super_raw_violin.pdf')
```

### Normalized

All observations have been scaled 
to the mean value of the wild-type strain of each dataset.
This looks good, but has the unfortunate side-effect
of zeroing the variance of the dataset means
in the wild-type strain.

This breaks all the assumptions of ANOVA 
if we wanted to run it on repeat-level means.
It will be better to show the raw data,
and make a linear model which accounts for batch effects
to get estimated means, effect sizes and p-values.

```{r}
plot_base <- ggplot(intensity_stats,
                   aes(x = strain, y = normalized_mean)) +
  labs(title = NULL, x = 'Strain', y = 'Normalized intensity (a. u.)')+
  scale_y_continuous(breaks = scales::breaks_extended(6))+
  scale_shape_manual(values = c(21:25)) +
  scale_color_brewer(palette = 'Set2') +
  scale_fill_brewer(palette = 'Set2') +
  facet_wrap(. ~localization,
             scales = 'free')


plot_scatter <- plot_base + 
  geom_quasirandom(inherit.aes = F, data = intensities,
                   aes(x = strain, y = normalized,
                       shape = dataset),
                   colour = 'gray75', #width = 0.3,
                   size = 0.8, #varwidth = TRUE,
                   show.legend = F
                   )


plot_violin <- plot_base + 
  geom_violin(inherit.aes = F, data = intensities,
              aes(x = strain, y = normalized),
              scale = 'width',
              colour = 'gray75', fill = 'transparent'
              )
```

```{r}
plot_super <- plot_scatter +
  geom_quasirandom(aes(shape = dataset, fill = dataset),
                   show.legend = F, width = 0.3, size = 2) +
  stat_summary(fun = mean, geom = "crossbar", 
               width = 0.5, fatten = 1) +
  stat_summary(fun.data = 'mean_sdl', geom = 'errorbar',
               fun.args = list(mult = 1), width = 0.2)
print(plot_super)
my_ggsave('figures/super_scatter.pdf')
```

```{r}
plot_super_violin <- plot_violin +
  geom_quasirandom(aes(shape = dataset, fill = dataset),
                   show.legend = F, width = 0.3, size = 2) +
  stat_summary(fun = mean, geom = "crossbar", 
               width = 0.5, fatten = 1) +
  stat_summary(fun.data = 'mean_sdl', geom = 'errorbar',
               fun.args = list(mult = 1), width = 0.2)
#print(plot_super_violin)
#my_ggsave('figures/super_violin.pdf')
```

```{r}
# These significance bars are generated by `ggsignif` from pairwise t-tests.
# They are not entirely correct and so they are not shown in the notebook.
# The figure is saved for me, to manually change the annotations
# in Illustrator to the ones from multiple comparisons tests, ie ANOVA.

plot_signif <- plot_super + 
  geom_signif(comparisons = list(c("wt", "3Δ"),
                                 c("3Δ", "oe"),
                                 c("wt", "oe")),
              map_signif_level = T, test = 't.test',
              step_increase = 0.03, tip_length = 0.01, vjust = 0.8)

my_ggsave('figures/plot_super_signif.pdf', plot = plot_signif)
```

## Modeling

We want to find out how the mean intensity values 
of cytoplasm or entire cells 
depend on the strain used 
(null hypothesis is that they do not).

In other experiments I followed the recommendations of
[several](https://doi.org/10.1083/jcb.202001064) 
[papers](https://doi.org/10.1091/mbc.E15-02-0076)
and performed hypothesis tests on repeat-level means.

This approach takes between-experiment variability into account,
and avoids p-value inflation, where tiny differences 
might end up 'highly significant' due to testing
a large number of observations which are not necessarily independent.

But it's also *very* conservative 
because using a small N of repeats means losing power
when compared to the hundreds of cell-level observations.

### Linear mixed model

Because we have a complete block design 
(all strains are represented in all batches),
we can include the dataset as a fixed or random effect.
Strain is obviously the fixed effect we want to analyze.
Because we do not care about estimating the effect
of any particular batch, a mixed model with the dataset
as a random effect will be appropriate.

#### Cellular

```{r}
cell_data <- intensities %>%
  filter(localization == 'Cellular')
```

Fitting a linear mixed model using the `lmerTest` package
(`lmer` comes from `lme4`, `lmerTest` just ads an interface for comparing models):

```{r}
cell_lmm <- lmer(intensity ~ strain + (1|dataset), data = cell_data)
summary(cell_lmm)
```

We have significant effect of `strain` on intensity
according to an F-test:

```{r}
anova(cell_lmm)
```

`dataset` does not seem to explain much variance,
but it definitely *could* 
due to the nature of measuring absolute fluorescence values,
as we see with the Cytosolic data.
It makes sense to keep it in the model.

In the plots we saw that maybe the data is not really normal,
as is typical for fluorescence values.
Model residuals are also not normal:

```{r}
qqnorm(resid(cell_lmm), main = 'LMM residuals Q-Q plot')
qqline(resid(cell_lmm))
hist(resid(cell_lmm), breaks = 100, ylab = 'Frequency', xlab = 'Residuals', main = 'LMM residuals histogram')
```

This is a *big* departure from normality, 
with a strong skew,
although it looks better on the histogram
More importantly, the variance in the data was also not homogeneous
(bigger spread in OE intensities).
We can see this reflected in the residuals:

```{r}
plot(cell_lmm, ylab = 'Residuals', xlab = 'Fitted values')
```

This is really not good, but then again 
the conclusions of the experiment are obvious
and do not hang on the quality of the modeling exercise.
Probably log-transform or a GLM could help;
lets stick to transforming the data.

We specify the LMM with `log(intensity) ~ strain + (1|dataset)`:

```{r}
cell_log_lmm <- lmer(log(intensity) ~ strain + (1|dataset), data = cell_data)
summary(cell_log_lmm)
```

We again have strong effect of `strain` and a random effect
with very little explanatory power for `dataset`.
Let's look at Q-Q plot and histogram of model residuals:

```{r}
qqnorm(resid(cell_log_lmm))
qqline(resid(cell_log_lmm))
hist(resid(cell_log_lmm), breaks = 100)
```

The residual distribution is still fat-tailed 
and skews slightly left now,
but the transformation helped.
More importantly, the differences in spread also decreased,
although they are still there:

```{r}
plot(cell_log_lmm, ylab = 'Residuals', xlab = 'Fitted values')
```

We could try a gamma GLM, but again,
the absolute quality of the model
is not critical to the conclusions here
so let's not over complicate.

I say we stop here and evaluate the group contrasts.
The `emmeans` package has a wonderful interface for pairwise contrasts,
*and* it computes group means with confidence intervals.

```{r}
cell_log_emm <- emmeans(cell_log_lmm, "strain")
```

The means and CIs are computed in log-space and back-transformed,
giving the absolute fluorescence values:

```{r}
summary(cell_log_emm, type = 'response') %>%
  tidy() %>%
  kable()
```

We can also compute differences between groups along with their CIs
and p-values for pairwise comparisons.
Because the differences are computed in log-space,
we get ratios after back-transforming.
This is actually very useful
because the ratio to wild-type
is what we wanted to know in the first place.

```{r}
summary(contrast(cell_log_emm, type = 'response', method = 'revpairwise'),
        infer = T) %>%
  kable()
```

Finally, the `emmeans` interface allows us to plot
the modeled means along with confidence intervals (bars)
and confidence intervals for the group differences (arrows).

```{r}
plot(cell_log_emm, comparisons = T, type = 'response', 
     xlab = 'Mean fluorescence +/- 95% CI (a.u.)')
```

We can conclude that all groups
are significantly different, 
with extremely low p-values (not that it matters).
3∆ background increases Ede1 levels by a modest ~20%,
and ADH1-driven overexpression by ~200%.

#### Cytosolic

```{r}
cyto_data <- intensities %>%
  filter(localization == 'Cytosolic')
```

After looking at the plots, 
we go into this exercise already suspecting 
that `strain` has no effect. But let's build a model as before:

```{r}
cyto_lmm <- lmer(intensity ~ strain + (1|dataset), data = cyto_data)
summary(cyto_lmm)
```

After modeling the fixed effect of `strain`,
dataset explains ~25% of the variance.
`strain` itself is not a significant predictor:

```{r}
anova(cyto_lmm)
```

We really could stop here, since `strain` does not explain anything,
and that's as clear from the model as from the plots.
But let's check model residuals:

```{r}
qqnorm(resid(cyto_lmm), main = 'Cytosolic LMM Q-Q plot')
qqline(resid(cyto_lmm))
plot(cyto_lmm, ylab = 'Residuals', xlab = 'Fitted values')
```

This is not so bad actually but we have some fat tails,
and a small difference in spread for larger values.
Log-transform helped before so let's see what it does here:

```{r}
cyto_log_lmm <- lmer(log(intensity) ~ strain + (1|dataset), data = cyto_data)
summary(cyto_log_lmm)
```

This is better, error-wise:

```{r}
qqnorm(resid(cyto_log_lmm), main = 'Cytosolic log_lmm Q-Q plot')
qqline(resid(cyto_log_lmm))
plot(cyto_log_lmm, ylab = 'Residuals', xlab = 'Fitted values')
```

But `strain` is still not significant:

```{r}
anova(cyto_log_lmm)
```

And if we use `emmeans` to get contrasts,
we find no significant differences between groups:

```{r}
cyto_log_emm <- emmeans(cyto_log_lmm, "strain")
summary(contrast(cyto_log_emm, type = 'response', method = 'revpairwise'),
        infer = T) %>%
  kable()
```

Since fluorescence intensity values are not meaningful,
I am skipping the estimated group means here.

## Conclusions

1. We cannot detect any differences 
    in cytosolic Ede1 intensities 
    across different strains.

2. We see significant differences in mean total cellular levels of Ede1, namely:
    * ~120% of wild-type Ede1 levels in 3∆ strains, and
    * ~320% of wild-type Ede1 levels in ADH1-driven overexpression.

A table summarising all group contrasts:

```{r}
tibble(summary(contrast(cyto_log_emm, type = 'response', method = 'revpairwise'),
        infer = T), localization = 'Cytosolic') %>%
  bind_rows(tibble(summary(contrast(cell_log_emm, type = 'response', method = 'revpairwise'),
        infer = T), localization = 'Cellular'))%>%
  select(contrast, ratio, lower.CL, upper.CL, localization) %>%
  mutate(across(where(is.numeric), round, digits = 2)) %>%
  kable()
```

## Source data

### .csv

```{r echo=FALSE}
xfun::embed_file('data/intensities.csv')
```

### .RData

```{r echo=FALSE}
xfun::embed_file('data/data_clean.RData')
```

## Session info

R session used to generate this document.

```{r session, message=TRUE}
sessionInfo()
```
