---
title: "Influence of Ede1 copy number on condensate formation in 3∆ diploid yeast"
date: "Last compiled on `r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    code_download: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE,
                      dpi = 96, fig.width = 6, fig.height = 4)
```

```{r libs}
library(tidyverse)
library(Hmisc) # mean_sdl and binomial confs
library(ggbeeswarm)
```

# {.tabset .tabset-pills}

## Experiment

The aim of this experiment was to quickly verify
whether the formation of Ede1 condensates
in living yeast cells
depends on Ede1 concentration.

I made three diploid strains homozygous for the
*apl3∆ yap1801∆ yap1802∆* triple adaptor deletion,
and differing in Ede1-EGFP copy number:
*EDE1-EGFP/EDE1/EGFP*, *EDE1-EGFP/EDE1*, *EDE1-EFGP/ede1∆*.

I took 5-micron z-stacks and manually cropped out single cells
using the brightfield channel as a reference to not bias the count.
A subjective, but blinded assessment was done 
for the presence / absence of condensates.
ImageJ macro `blindAnalysis` was used 
to display all cells in random order
and with hidden strain information,
and to record the binary assessment.

### Strain list

```{r strains}
strains <- read_csv("data/raw/strain_lut.csv", 
                    col_types = cols(second_ede1  = col_factor(NULL))) # order factor levels in csv file
knitr::kable(strains)
```

## Results {.tabset}

```{r source_data}
# initialize a data frame with counts and get information about experiment from file names
files <- list.files(path = 'data/raw/', pattern = 'dataset',
                    full.names = TRUE)

counts <- map(files, read_csv,
              col_names = c('cell', 'condensates', 'dataset'),
              cols(dataset = col_factor(NULL))) %>%
          bind_rows() %>%
          separate(cell, c('date', 'strain', 'stack', 'cell',
                   'channel'), '_') %>%
          left_join(strains, by = 'strain') %>%
          select(strain, stack, second_ede1, condensates, dataset)

write_excel_csv(format.data.frame(counts, digits = 3), 'data/diploid_data.csv')
save(counts, file = 'data/diploid_data.RData')
rm(list = ls()) # clean workspace
```

```{r load}
load('data/diploid_data.RData')
```

### Individual datasets

Percentage of cells with condensates 
for each EDE1-EGFP / second Ede1 allele combination,
by dataset.

x: number of cells with condensates, 
n: total number of analyzed condensates, 
p: x / n, 
p_min and p_max: estimates of binomial confidence interval 
based on sample size, by Wilson's method.

```{r individual_datasets}
proc <- counts %>%
  group_by(second_ede1, dataset) %>%
  summarise(x = sum(condensates), n = n(),
            p = binconf(x, n, method = 'wilson')[1],
            p_min = binconf(x, n, method = 'wilson')[2],
            p_max = binconf(x, n, method = 'wilson')[3])

proc_stats <- proc %>%
  select(second_ede1, dataset, p) %>%
  group_by(second_ede1) %>%
  summarise(mean = mean(p), sd = sd(p), n = n())

knitr::kable(proc)
write_excel_csv(format.data.frame(proc, digits = 3), 'data_summary.csv')
```

### Averaged datasets

Percentage of cells bearing condensates for each
EDE1-EGFP / second Ede1 allelecombination,
arithmetic mean ± SD from 3 datasets.

```{r averaged}
proc_stats <- proc %>%
  select(second_ede1, dataset, p) %>%
  group_by(second_ede1) %>%
  summarise(mean = mean(p), sd = sd(p), n = n())

knitr::kable(proc_stats)
```

### Plot {.active}

```{r plot}
theme_frap <- function(base_size = 11, base_family = "",
                       base_line_size = base_size / 22,
                       base_rect_size = base_size / 22) {
  # minimal theme with border
  # based on theme_linedraw without the grid lines
  # also trying to remove all backgrounds but plot_background won't set to blank
  theme_linedraw(
    base_size = base_size,
    base_family = base_family,
    base_line_size = base_line_size,
    base_rect_size = base_rect_size
  ) %+replace%
    theme(
      # no grid and no backgrounds if I can help it
      plot.background = element_blank(),
      panel.background = element_blank(),
      panel.grid = element_blank(),
      plot.margin = margin(0, 0, 0, 0),
      complete = TRUE
    )
}

myplot <- proc %>% ggplot(aes(x = second_ede1, y = p)) + 
  geom_quasirandom(aes(shape = dataset), fill = 'gray80',
              size = 2, width = 0.3, show.legend = FALSE)+
  stat_summary(fun.y = mean, geom = "errorbar", 
               aes(ymax = ..y.., ymin = ..y..),
               width = 0.5, size = 1)+
  stat_summary(fun.data = 'mean_sdl', geom = 'errorbar',
               fun.args = list(mult = 1), width = 0.2)+
  #scale_color_brewer(palette = 'Set2')+
  scale_shape_manual(values = c(21:25))+
  theme_frap(base_size = 18, base_family = "Myriad Pro")+
  scale_y_continuous(labels = scales::label_percent(suffix = ''))+
  theme(legend.position = c(0.95, 0.95),
        legend.justification = c("right", "top"),
        legend.background = element_rect(fill = NULL, colour = 'black',
                                         size = 0.5))+
  labs(title = NULL,
       y = 'Cells with condensates (%)',
       x = 'Second Ede1 allele')

print(myplot)
ggsave('figures/diploids.pdf', device = cairo_pdf, width = 80, height = 120, units = "mm")
```

## Source data

### .csv

```{r echo=FALSE}
xfun::embed_file('data/diploid_data.csv')
```

### .RData

```{r echo=FALSE}
xfun::embed_file('data/diploid_data.RData')
```

## Session info

```{r session, message=TRUE}
sessionInfo()
```
